### 2020-06-25 Documents/ESA/sections/ProposedResearch.tex
minor edits throughout the document

	\begin{itemize}
		\item \textbf{Stability and Chaos in MARL:} in which we seek to understand the dynamics of
		agent strategies when using Q-Learning to learn a game through iteration. We will establish
		the stability, as a function of parameters, when learning general p-player N-action games.
		This study enables the appropriate selection of parameters and payoff matrices to ensure the
		stability of the strategies of a finite set of agents, such as leaders in a swarm.
		\item \textbf{Dynamics of Mean-Field Q-Learning Games:} in which we examine the
		strategy dynamics for large populations of agents learning through iterated games and
		mean-field Q-Learning. We seek to understand the long term behaviour of such mean-field
		systems in terms of its strategy selection. This study extends the previous and allows for
		the stability of the strategies of a population of agents to be established.
		\item \textbf{Model Predictive Control of Active Particles through Fields:} in which we
		investigate the interaction of a swarm of active particles with potential fields. We
		establish the stabilisability of an MPC scheme with defined stage costs, as well as an
		analysis of the suboptimality of the method.
		\item \textbf{Incorporation of Intelligence in Control:} in which we adapt the dynamical
		system from the above point to include an interaction term, accounting for strategy
		selection by agents who learn through iterated games against one another. Interactions with
		an MPC scheme is then to be examined in a similar capacity.

@@

@@

@@

@@

    One of the main results shown by Sanders et al. \cite{Sanders2018} is that as the number of
    players in a game increases, the learning behaviour is more likely to be chaotic, regardless of
    the choice of parameters. This is intuitive since a higher number of players would result in a
    greater strategy space and more agents for any particular player to learn against and is
    verified by their presented results as in Figure \ref{fig::GallaPrevalence} - the hotter
    regions occupy a larger area as

@@

    Yet it can be argued that, for large populations of agents (e.g. a crowd), the aggregate
    behaviour may be predictable. This intuition is the foundation upon which crowd dynamics and
    flocking systems are based. In fact, the work presented by Leung et al. \cite{Hu2019} provides a
    cursory verification of this intuition. Here, the authors present an anaysis of the learning
    dynamics for a large agent population (which they approximate as containing infinite agents)
    where each agent is an independent Q-Learning using Boltzmann action selection. The
    result is a system of equations, governed by a Fokker-Planck model which is numerical shown to be a 
    strong predictor of the overall strategy selection of the population. Indeed, rather than
    examining the strategies of every agent in the population (which the authors approximate as
    infinitely large), they model the evolution of a probability density function which, for each
    action, defines the density of agents who retain a given Q-value for that action. This is
    illustrated in Figure \ref{fig::LeungPredictions}.

@@

    This section of the study is perhaps the strongest extension proposed in this
	chapter, and will likely be the most challenging. Here, we examine the dynamics (
	\ref{eqn::Vlasov}) in which the interaction term accounts for the social dynamics of the
	decision making of individual agents who have learnt through an iterated game. We leverage the
	strategy evolution dynamics and stability analysis established in 
	\ref{sec::Chaos_in_MARL} and \ref{sec::Large_Agent_Dynamics}.

>>>>

	\begin{itemize}
		\item \textbf{Stability and Chaos in MARL:} in which
                  we seek to understand the dynamics of agent
                  strategies when using Q-Learning to learn a game
                  through iteration. We will establish the stability,
                  as a function of parameters, when learning general
                  p-player N-action games.  This study enables the
                  appropriate selection of parameters and payoff
                  matrices to ensure the stability of the strategies
                  of a finite set of agents, such as leaders in a
                  swarm.

                  \fr{this is already attempted in the currect
                    research section, isn't it?}

              \item \textbf{Dynamics of Mean-Field Q-Learning Games:} in which we examine the
		strategy dynamics for large populations of agents learning through iterated games and
		mean-field Q-Learning. We seek to understand the long term behaviour of such mean-field
		systems in terms of its strategy selection. This study extends the previous and allows for
		the stability of the strategies of a population of agents to be established.
		\item \textbf{Model Predictive Control of Active Particles through Fields:} in which we
		investigate the interaction of a swarm of active particles with potential fields. We
		establish the stabilisability of an MPC scheme with defined stage costs, as well as an
		analysis of the suboptimality of the method.
		\item \textbf{Incorporation of Intelligence in Control:} in which we adapt the dynamical
		system from the above point to include an interaction term, accounting for strategy
		selection by agents who learn through iterated games against one another. Interactions with
		an MPC scheme is then to be examined in a similar capacity.

@@

        \fr{we analyze each topic in a section \ldots}

@@

        \fr{I understand the equations are the same as in the previous section. I suggest we move the explanation there and remove it from here.}

@@

    \fr{I would move all of the above to the previous section.}

@@

    One of the main results shown by Sanders et al. \cite{Sanders2018}
    is that as the number of players in a game increases, the learning
    behaviour is more likely to be chaotic, regardless of the choice
    of parameters. This is intuitive since a higher number of players
    would result in a greater strategy space and more agents for any
    particular player to learn against and is verified by their
    presented results as in Figure \ref{fig::GallaPrevalence} - the
    hotter regions occupy a larger area as \fr{?}

@@

    Yet it can be argued that, for large populations of agents (e.g.,
    a crowd), the aggregate behaviour may be predictable \fr{provide some justification to this claim?}. This
    intuition is the foundation upon which crowd dynamics and flocking
    systems are based. In fact, the work presented by Leung et
    al.~\cite{Hu2019} provides a cursory verification of this
    intuition. Here, the authors present an analysis of the learning
    dynamics for a large agent population (which they approximate as
    containing infinite agents) where each agent is an independent
    Q-Learning using Boltzmann action selection. The result is a
    system of equations, governed by a Fokker-Planck model which is
    numerical shown to be a strong predictor of the overall strategy
    selection of the population. Indeed, rather than examining the
    strategies of every agent in the population (which the authors
    approximate as infinitely large), they model the evolution of a
    probability density function which, for each action, defines the
    density of agents who retain a given Q-value for that action. This
    is illustrated in Figure \ref{fig::LeungPredictions}.

@@

    This section of the study is perhaps the strongest extension
    proposed in this chapter, and will likely be the most
    challenging. Here, we examine the dynamics (\ref{eqn::Vlasov}) in
    which the interaction term accounts for the social dynamics of the
    decision making of individual agents who have learnt through an
    iterated game. We leverage the strategy evolution dynamics and
    stability analysis established in \ref{sec::Chaos_in_MARL} and
    \ref{sec::Large_Agent_Dynamics}.

### 2020-06-25 Documents/ESA/sections/ShortTermGoals.tex
minor edits throughout the document

    In this chapter, we summarise and present the research and results which have been obtained thus
    far. These tackle the problems described in Section \ref{sec::Chaos_in_MARL}. We detail the
    derivation of a stability phase line as presented in Sanders et al. \cite{Sanders2018} as well
    as the numerical simulations performed to verify these results. The steps to be completed
    subseqent to this study are then given.

@@

    We begin by considering the evolution of learning with agents who follow a Q-Learning approach.
    Specifically, we look at characterising the stability of agents learning strategies. This
    technique is carried out by Sanders et al \cite{Sanders2018}, in which the authors
    characterise the strategy evolution of agents who learn using an 'Experience Weighted
    Attraction' (EWA) algorithm, which is seen to be a strong representation of how people learn in
    games. The authors are able to determine the regions of parameter space in which learning is
    likely to converge towards stable equilibriums, as opposed to complex behaviours (such as limit
    cycles or chaotic dynamics). The aim of this study is to format these techniques for the study
    of computational agents who follow the popular Q-Learning approach \cite{Sutton2018,SchwartzMulti-agentApproach}.

@@

    With the accuracy of the continuous time dynamical system (\ref{eqn::EOM}) established, we analyse the
    stability of this system in the following section. Note that whilst the derivation provided is
    for the particular case of two agents for the sake of brevity, the problem of p-player games is
    equivalent to that of two players and so the solutions to the former can (and will) be
    presented.

@@

     We begin with the equations as described by Tuyls et al \cite{Tuyls2006AnGames}

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders et al. refer to these as the memory and intensity of choice parameters respectively. Agent 1 takes action i with probability $x_i$ while Agent 2 takes action j with probability $y_j$. If these actions are taken, the agents receive payoff $a_{ij}$ and $b_{ji}$ respectively.

    Our goal is to analyse the dynamics of any game, regardless of payoff elements. As such, we
    follow the approach of Sanders et al. in deriving the full set of 'effective dynamics' averaged
    over all possible realisations of the payoff matrices. Note that these elements are drawn from a
    Multivariate Gaussian such that

    \begin{equation*}
        \begin{split}
            \mathbb{E}[a_{ij}] = \mathbb{E}[b_{ji}] = 0\\
            \mathbb{E}[a_{ij}^2] = \mathbb{E}[b_{ji}^2] = 1\\
            \mathbb{E}[a_{ij} b_{ji}] = \Gamma.
        \end{split}
    \end{equation*}

    where $\Gamma$ is a parameter which defines the 'competitiveness' of the game. For a two player
    game, $\Gamma \in [-1, 1]$ where -1 denotes a zero-sum game and 1 denotes a game with positive
    correlation across payoff elements. Taking this average yields

    \begin{equation}
    \begin{split}
        \dot{x}(t) &= x(t)(\Gamma \alpha^2 \tilde{\tau}^2 \int dt'[G_y(t, t')x(t')] + 
        \tilde{\alpha}
        \rho_x(t) + \alpha \tilde{\tau} \eta_x(t) + \tilde{\alpha} \tau \eta_x(t) \eta_y(t) +
        \sqrt{\Gamma} \tilde{\alpha} \tau \mu_x) \\
        \dot{y}(t) &= y(t)(\Gamma \alpha^2 \tilde{\tau}^2 \int dt'[G_x(t, t')y(t')] + 
        \tilde{\alpha} \rho_y(t) +
        \alpha \tilde{\tau} \eta_y(t) + \tilde{\alpha} \tau \eta_x(t) \eta_y(t)+ 
        \sqrt{\Gamma} \tilde{\alpha} \tau \mu_y), \\
    \end{split}
    \end{equation}

    Here, $G, \rho, \eta, \mu$ are correlation functions, generated due to the averaging process 
    (see Appendix \ref{app::ResearchSummary}). We now analyse the stability of this dynamical system
    by linearising about
    a
    fixed point. This yields the requirement that a fixed point $x^*$ satisfy

    \begin{equation}
    0 = x^* \left[\Gamma \alpha^2 \tilde{\tau}^2 x^* \int G_y(t - t') dt' + \tilde{\alpha} \rho_x^* + \alpha \tilde{\tau} \eta_x^* + \tilde{\alpha} \tau \eta_x^* \eta_y^* + \sqrt{\Gamma} \tilde{\alpha} \tau \mu_x \right].
    \end{equation}

    The implication here is that the resulting value of $x^*$ can take two solutions. However, one of these is $x^* = 0$, which is found to rarely occur, while the second solution contains a term $\sqrt{\Gamma}$, which would yield complex values. If this is true, then it would imply that the learning algorithm will rarely converge for values of $\Gamma < 0$. Numerical experiments will shed more light on the likelihood of convergence. If this ansatz turns out to be incorrect (i.e. we see significant convergence for $\Gamma < 0$), it is unlikely that we will be able to solve for the stability line for this case.

    We take the Fourier transform of this system to determine the long term behaviour of thhe system
    after a disturbance. To neglect all transient behaviour, we consider the equation at $\omega =
    0$

@@

\begin{enumerate}
    \item Fix the parameters $\alpha$ and $\gamma$. The latter is held fixed at $0.1$ since
    it does not affect the long term behaviour of the system (it does not appear in (
    \ref{eqn::EOM})).
    \item We initialise values of $\Gamma$ and $\tau$. These will be the variables which we sweep
    over.
    \item Generate payoff matrices for both agents by sampling from a multi-variate Gaussian 
    (variables are the payoff elements) with covariance parameterised by $\Gamma$.
    \item Iniitialise the agents with random initial conditions (i.e. random action probabilities).
    \item Allow the agents to learn over a maximum of $1 \times 10^5$ iterations.
    \item Every 100 iterations, check to see if the action probabilities have changed significantly.
    If not (i.e. the change over the last 100 iterations is less than 1\%) the learning is
    considered to have converged.
    \item This process is repeated 5 times with random payoff matrices generated based on the value
    of $\Gamma$. The probability of convergence is then recorded as $\frac{\text{number of times
    converged}}{5}$.
    \item The values of $\tau$ and $\Gamma$ are then modified and the process is repeated. The
    heatmap shows the probability of convergence for all values of $\tau$ and $\Gamma$ which are
    tested.
\end{enumerate}

@@

    We then seek to expand this study to a population of agents. For this, we will study the
    stability of the system presented by Leung et al \cite{Hu2019} which presents a mean-field model
    describing the evolution of learning of a population of Q-Learning agents who play co-operative
    games against one another. This last point is an important limitation - the derived dynamics
    will only consider populations who must co-operate with one another (i.e. the $\Gamma$ value of
    the game is one).

>>>>

    In this chapter, we summarise and present the research and results
    which have been obtained thus far. These tackle the problems
    described in Section \ref{sec::Chaos_in_MARL}. We detail the
    derivation of a stability phase line as presented in Sanders et
    al.~\cite{Sanders2018} as well as the numerical simulations
    performed to verify these results. The steps to be completed
    subseqent to this study are then given.

\fr{perhaps we can provide a bit more of context. Also, we can refer to the appendix for the details of the derivation and mention that we are aiming for a submission to AAAI.}

@@

    We begin by considering the evolution of learning with agents who
    follow a Q-Learning approach.  Specifically, we look at
    characterising the stability of agents learning strategies. This
    technique is carried out by Sanders et al.~\cite{Sanders2018}, in
    which the authors characterise the strategy evolution of agents
    who learn using an 'Experience Weighted Attraction' (EWA)
    algorithm \cite{?}, which is seen to be a strong \fr{why?} representation of how
    people learn in games. The authors are able to determine the
    regions of parameter space in which learning is likely to converge
    towards stable equilibriums, as opposed to complex behaviours
    (such as limit cycles or chaotic dynamics). The aim of this study
    is to format these techniques for the study of computational
    agents who follow the popular Q-Learning approach
    \cite{Sutton2018,SchwartzMulti-agentApproach}.

@@

    With the accuracy of the continuous time dynamical system
    (\ref{eqn::EOM}) established, we analyse the stability of this
    system in the following section. Note that whilst the derivation
    provided is for the particular case of two agents for the sake of
    brevity, the problem of $p$-player games is equivalent to that of
    two players and so the solutions to the former can (and will) be
    presented.

    \fr{differences/similarities between EWA and Q-learning?}

@@

     We begin with the equations as described by Tuyls et al \cite{Tuyls2006AnGames} \fr{describing \ldots}

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders
    et al. refer to these as the {\em memory} and {\em intensity of choice}
    parameters respectively. Agent 1 takes action $i$ with probability
    $x_i$ while Agent 2 takes action $j$ with probability $y_j$. If
    these actions are taken, the agents receive payoff $a_{ij}$ and
    $b_{ji}$ respectively.

    Our goal is to analyse the dynamics of any game, regardless of
    payoff structure. As such, we follow the approach of Sanders et
    al. in deriving the full set of 'effective dynamics' averaged over
    all possible realisations of the payoff matrices. Note that these
    elements are drawn from a Multivariate Gaussian such that
    \begin{equation*}
        \begin{split}
            \mathbb{E}[a_{ij}] = \mathbb{E}[b_{ji}] = 0\\
            \mathbb{E}[a_{ij}^2] = \mathbb{E}[b_{ji}^2] = 1\\
            \mathbb{E}[a_{ij} b_{ji}] = \Gamma.
        \end{split}
    \end{equation*}
    where $\Gamma$ is a parameter which defines the `competitiveness' of the game. For a two player
    game, $\Gamma \in [-1, 1]$ where -1 denotes a zero-sum game and 1 denotes a game with positive
    correlation across payoff elements. Taking this average \fr{some more details?} yields
%
    \begin{equation}
    \begin{split}
        \dot{x}(t) &= x(t)(\Gamma \alpha^2 \tilde{\tau}^2 \int dt'[G_y(t, t')x(t')] + 
        \tilde{\alpha}
        \rho_x(t) + \alpha \tilde{\tau} \eta_x(t) + \tilde{\alpha} \tau \eta_x(t) \eta_y(t) +
        \sqrt{\Gamma} \tilde{\alpha} \tau \mu_x) \\
        \dot{y}(t) &= y(t)(\Gamma \alpha^2 \tilde{\tau}^2 \int dt'[G_x(t, t')y(t')] + 
        \tilde{\alpha} \rho_y(t) +
        \alpha \tilde{\tau} \eta_y(t) + \tilde{\alpha} \tau \eta_x(t) \eta_y(t)+ 
        \sqrt{\Gamma} \tilde{\alpha} \tau \mu_y), \\
    \end{split}
    \end{equation}

    Here, $G, \rho, \eta, \mu$ are correlation functions, generated
    due to the averaging process (see Appendix
    \ref{app::ResearchSummary}). We now analyse the stability of this
    dynamical system by linearising about a fixed point. This yields
    the requirement that a fixed point $x^*$ satisfy
%
    \begin{equation}
    0 = x^* \left[\Gamma \alpha^2 \tilde{\tau}^2 x^* \int G_y(t - t') dt' + \tilde{\alpha} \rho_x^* + \alpha \tilde{\tau} \eta_x^* + \tilde{\alpha} \tau \eta_x^* \eta_y^* + \sqrt{\Gamma} \tilde{\alpha} \tau \mu_x \right].
    \end{equation}

    The implication here is that the resulting value of $x^*$ can take two solutions. However, one of these is $x^* = 0$, which is found to rarely occur, while the second solution contains a term $\sqrt{\Gamma}$, which would yield complex values. If this is true, then it would imply that the learning algorithm will rarely converge for values of $\Gamma < 0$. Numerical experiments will shed more light on the likelihood of convergence. If this ansatz turns out to be incorrect (i.e., we see significant convergence for $\Gamma < 0$), it is unlikely that we will be able to solve for the stability line for this case.

    We take the Fourier transform of this system \ref{} to determine the long term behaviour of the system
    after a disturbance. To neglect all transient behaviour, we consider the equation at $\omega =
    0$

@@

\begin{enumerate}
    \item Fix the parameters $\alpha$ and $\gamma$. The latter is held fixed at $0.1$ since
    it does not affect the long term behaviour of the system (it does not appear in (
    \ref{eqn::EOM})).
    \item We initialise values of $\Gamma$ and $\tau$. These will be the variables which we sweep
    over.
    \item Generate payoff matrices for both agents by sampling from a multi-variate Gaussian 
    (variables are the payoff elements) with covariance parameterised by $\Gamma$.
    \item Initialise the agents with random initial conditions
      (i.e., random action probabilities).
    \item Allow the agents to learn over a maximum of $1 \times 10^5$ iterations.
    \item Every 100 iterations, check to see if the action
      probabilities have changed significantly.  If not (i.e., the
      change over the last 100 iterations is less than 1\%) the
      learning process is considered to have converged.

    \item This process is repeated 5 times with random payoff matrices generated based on the value
    of $\Gamma$. The probability of convergence is then recorded as $\frac{\text{number of times
    converged}}{5}$.

  \item The values of $\tau$ and $\Gamma$ are then modified and the process is repeated. The
    heatmap shows the probability of convergence for all values of $\tau$ and $\Gamma$ which are
    tested.
\end{enumerate}

@@

    We then seek to expand this study to a population of agents. For
    this, we will study the stability of the system presented by Leung
    et al \cite{Hu2019} which presents a mean-field model describing
    the evolution of learning of a population of Q-Learning agents who
    play co-operative games against one another. This last point is an
    important limitation - the derived dynamics will only consider
    populations who must co-operate with one another (i.e., the
    $\Gamma$ value of the game is 1).
